---
title: "Append to envirodata"
output: html_document
date: "2023-11-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load libraries
```{r}
source("PA-paper/LoadPackages_PA.R")
library(sf)
```

bring in and format presences and pseudo-absences
```{r}
presence<-read.csv("20231129SSMoutput.csv")

# List all files in the directory
files <- list.files("PA-paper/output.Predicted")
csv_files <- files[grepl("\\.csv$", files)]

# Initialize an empty data frames
crw_data <- data.frame()
background_data <- data.frame()

# Loop through each file
for (file in csv_files) {
  
  # Extract the tag from the file name
  tag <- gsub(".*sim_|.csv.*", "", file) #start here
  cat("doing", file, "n/")
  # Check the file name to determine the data type
  if (grepl("^crw_sim", file)) {
    data_type <- "crw"
  } else if (grepl("^backgroundpts_sim", file)) {
    data_type <- "background"
  } else {
    stop("Invalid file name:", file)
  }

  # Extract the tag from the file name
  tag <- gsub(".*sim_|.csv.*", "", file) #start here

  # Read the data from the file
  file_data <- read.csv(paste("PA-paper/output.Predicted/", file, sep=""))

  # Add columns for data type and tag
  file_data$data_type <- data_type
  file_data$tag <- tag

  # Add data to the appropriate data frame
  if (data_type == "crw") {
    crw_data <- rbind(crw_data, file_data)
  } else {
    background_data <- rbind(background_data, file_data)
  }
}

#write.csv(background_data, file="PA-paper/background.csv")
#write.csv(crw_data, file="PA-paper/crw_data.csv")
```
Combine presences, crw and background data
```{r}
crw_data<-read.csv("PA-paper/crw_data.csv")
background_data<-read.csv("PA-paper/background.csv")

presence1<-presence |> dplyr::select(c(id, lon, lat, date))
presence1$data_type<-"track"
presence1$iteration<-NA
presence1<-presence1 |> mutate(date=as.POSIXct(date, format="%Y-%m-%d %H:%M:%S"))

background_data<-background_data |> 
  mutate(id=tag, date=as.POSIXct(dTime, format="%Y-%m-%d %H:%M:%S")) |> 
  dplyr::select(!c(X,dTime, step, ID, optional, tag))

crw_data<-crw_data |> 
  mutate(id=tag, lon=x, lat=y, date=as.POSIXct(t, format="%Y-%m-%d %H:%M:%S")) |> 
  dplyr::select(!c(X,tag, flag, t, x, y))

all<-rbind(presence1, background_data, crw_data)
all$date1<-as.POSIXct(all$date, format="%Y-%m-%d %H:%M:%S")
#write.csv(all, file="20231206combinedPA.csv")
```

# Extracting spatial and spatio-temporal environmental covariates

1. Static variables
Sampled predictor layers (creation/modification of source datasets specified in Table 1 of the paper)
Created a mask layer from the extent of the Carter et al. dataset ("mask.tif")
Clipped input datasets to this extent

QGIS a) sample raster tool for grid data, and b)join attributes by location tool for shapefile (categorical) data

for b) features contain or overlap points. Join type one to many.

## Shapefiles
- built features
- geohabitat
- geomorphology
- seagrass community type


## Rasters
- bathymetry
- distance to rivers 
- distance to reefs
- distance to coast
- distance to rec boats
- seagrass probability
- relative tidal exposure
- MMP - frequency of exposure to WT1 and WT2 2003-2022
- MMP - longterm exposure
- ruggedness
- slope


2. DEA mangrove extraction
```{r}
#Extracted yearly raster via DEA sandbox environment
#Filtered for values 1-4 in qgis
#Proximity rasters in qgis
#Append per matching year:
# pts<-nas%>%mutate(date1=as.Date(aus.date))
# clipped<-read.csv("PA-paper/crw_data.csv")
# 
# df2_selected<-clipped %>%
#   rename(lat=y,
#          lon=x,
#          id=tag,
#          date=t)%>%
#   dplyr::select(id, lon, lat, data_type, iteration, date)%>%
#   mutate(date1=as.Date(date))
# filtered<-pts%>%
#   left_join(df2_selected, by=c("lat", "lon", "id", "data_type", "iteration", "date1"))
# 
# #convert timestamp to year
# filtered$date<-as.POSIXct(filtered$date.y, format="%Y-%m-%d %H:%M:%S")
# filtered<-filtered |> mutate(aus.date=date+hours(10))
# points<-filtered%>%mutate(aus.date=as.Date(aus.date))

# points$year <- format(as.Date(points$aus.date), "%Y")
# 
# 
# pts_sf<-st_as_sf(points, coords=c("lon", "lat"), crs=4326)
# 
# 
# raster_list<-lapply(2010:2022, function(year) raster(paste0("enviro datasets/DEA mangroves/Reprojected", year, ".tif")))
```


```{r}
# pts_sf$deamangrove<-NA
# 
# for(i in 1:length(raster_list)){
#   year <- 2010 + i - 1  # Calculate the corresponding year
#   cat("doing", year)
#   matching<-which(pts_sf$year==year)
#   if(length(matching)>0){
#     # Extract raster values for matching points
#     pts_sf[matching,]$deamangrove <- raster::extract(raster_list[[i]], st_coordinates(pts_sf[matching, ]), method='simple')
# 
#     # Assign extracted values to the raster_value column in points
# 
#   }
# }
```


4.eReefs:
EFI, wind speed, current, salinity, Secchi depth, temperature, zooplankton N
```{r}
#Query AIMS thredds server via ExtractEReefsCRW.R and ExtractEReefsbackground.R
# break up dataset for processing
clipped<-read.csv("20240110clippedwrasters.csv")
track<-clipped |> filter(data_type=="track")
#write.csv(track, "20240110trackwrasters.csv")
crw<-clipped |> filter(data_type=="crw")
#write.csv(crw, "20240110crwwrasters.csv")
background<-clipped |> filter(data_type=="background")
#write.csv(background, "20240110backgroundwrasters.csv")
```

Gluing the outputs of the eReefs query scripts back together:
```{r}
# hyd4<-read.csv("hydro_4dimscrw.csv")
# hyd3<-read.csv("hydro_3dimscrw.csv")
# bgc4<-read.csv("bgc_4dimscrw.csv")
# bgc3<-read.csv("bgc_3dimscrw.csv")
# 
# 
# # arrange dfs by index
# hyd4test<-hyd4%>%arrange(X.3)

# # add in hyd3dims vars
# 
# hyd3vars<-hyd3%>%arrange(X.3)
# hydtest<-cbind(hyd4test, hyd3vars[, c("mean_wspeed")])
# 
# # add in bgc vars
# bgc4vars<-bgc4%>%arrange(X.3)
# bgc3vars<-bgc3%>%arrange(X.3)
# bgc4test<-cbind(hydtest, bgc4vars[, c("ZooL_N", "EFI")])
# bgc3test<-cbind(bgc4test, bgc3vars[, c("Secchi")])
# bgc3test<-bgc3test%>%rename("mean_wspeed"=`hyd3vars[, c("mean_wspeed")]`, 
#                             "Secchi"=`bgc3vars[, c("Secchi")]`)
# # save this 
# write.csv(bgc3test, "20240304crw.csv")
# Clipped to seagrass raster data- resulting in:
# crw<-read.csv("20240304crwwrug.csv")
```


##########################################

# For grid prediction:
## Convert polygon layers to rasters
```{r}
COMMUNITY<-terra::vect("enviro datasets/NESP predicted distribution of seagrass communities/Shapefiles/GBR_NESP-TWQ-5.4_JCU_Seagrass-communities_20201203.shp")
COMMUNITY$COMMUNITY<-as.factor(COMMUNITY$COMMUNITY)
y=terra::rast(COMMUNITY, resolution=common_res, extent=common_extent, crs=common_crs)
seagrassCom<-terra::rasterize(COMMUNITY, y, field='COMMUNITY')
seagrasscrop <- crop(seagrassCom, extent(extent))
plot(seagrasscrop)
ra<-aggregate(seagrasscrop, fact=2)
raster::writeRaster(ra, "COMMUNITY.asc")

builtfeature<-terra::vect("enviro datasets/Breakwaters, groynes, seawalls, marinas/QSC_Extracted_Data_20231018_085138072000-40848/data.gdb")
builtfeature$FEATURETYPE<-as.factor(builtfeature$FEATURETYPE)
y=terra::rast(builtfeature, resolution=common_res, extent=common_extent, crs=common_crs)
builtfeature<-terra::rasterize(builtfeature, y, field='FEATURETYPE')
builtfeatcrop <- crop(builtfeature, extent(extent))
plot(builtfeatcrop)
ra<-aggregate(builtfeatcrop, fact=2)
raster::writeRaster(ra, "builtfeature.asc")


geohab<-terra::vect("enviro datasets/qld_geohab_av/czm/pristine_ests/release/qld/shape/geohab_qld_v2.shp")
geohab$GH_TYPE<-as.factor(geohab$GH_TYPE)
y=terra::rast(geohab, resolution=common_res, extent=common_extent, crs=common_crs)
geohab<-terra::rasterize(geohab, y, field='GH_TYPE')
geohabcrop <- crop(geohab, extent(extent))
plot(geohabcrop)
ra<-aggregate(geohabcrop, fact=2)
raster::writeRaster(ra, "geohab.asc")

marina<-terra::vect("enviro datasets/Breakwaters, groynes, seawalls, marinas/QSC_Extracted_Data_20231018_085120960000-49924/data.gdb")
marina$FEATURETYPE<-as.factor(marina$FEATURETYPE)
y=terra::rast(marina, resolution=common_res, extent=common_extent, crs=common_crs)
marina<-terra::rasterize(marina, y, field='FEATURETYPE')
marinacrop <- crop(marina, extent(extent))
plot(marinacrop)
ra<-aggregate(marinacrop, fact=2)
raster::writeRaster(ra, "marina.asc")

recboating<-terra::vect("enviro datasets/rec_boating_buffered.shp")
recboating$FACILITY<-as.factor(recboating$FACILITY)
y=terra::rast(recboating, resolution=common_res, extent=common_extent, crs=common_crs)
recboating<-terra::rasterize(recboating, y, field='FACILITY')
recboatingcrop <- crop(recboating, extent(extent))
plot(recboatingcrop)
ra<-aggregate(recboatingcrop, fact=2)
raster::writeRaster(ra, "recboating.asc")

geomorph<-terra::vect("enviro datasets/Heap2008Geomorphology.shp")
geomorph$feature<-as.factor(geomorph$feature)
y=terra::rast(geomorph, resolution=common_res, extent=common_extent, crs=common_crs)
geomorph<-terra::rasterize(geomorph, y, field='feature')
geomorphcrop <- crop(geomorph, extent(extent))
plot(geomorphcrop)
ra<-aggregate(geomorphcrop, fact=2)
raster::writeRaster(ra, "geomorph.asc")

```
## real time ereefs layers as rasters
```{r}
nc_file <- "enviro datasets/ascii rasters for predictions/gbr4_simple_2023-10-19.nc"
nc_file_bgc <- "enviro datasets/ascii rasters for predictions/gbr4_bgc_simple_2023-10-19.nc"
```
```{r}
variables_hydro <- c("temp", "salt", "v", "u", "mean_cur", "mean_wspeed", "wspeed_u", "wspeed_v")  # List your variables here
variables_bgc <-c("Secchi", "EFI", "ZooL_N")
```
```{r}
raster_layer <- raster(nc_file, varname = 'wspeed_u')
nc_data<-nc_open(nc_file)
var_data <- ncvar_get(nc_data, variables_hydro[1])
var_data_permuted <- aperm(var_data, c(1:(dim(var_data)-2), dim(var_data)[2], dim(var_data)[1]))
plot(raster_layer)
```

```{r}
for (var in variables_hydro) {
  # Read the NetCDF variable as a raster
  r <- raster(nc_file, varname = var)
  
  # (Optional) Re-grid - create a target raster for desired resolution and extent
  target_raster <- raster(extent(r), res=common_res)
  r_resampled <- resample(r, target_raster, method="bilinear")
  
  # Clip to the shapefile
  r_clipped <- mask(r_resampled, shp)
  
  # Set CRS
  crs(r_clipped) <- CRS(common_crs)
  
  # Save as ASCII raster
  ascii_filename <- paste0(var, "2023.asc")  # Construct filename
  writeRaster(r_clipped, ascii_filename, format="ascii")
}
```
