---
title: "presence and pseudo-absence setup"
output: html_document
date: "2023-08-16"
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(stars)
library(readr)
library(SDLfilter)
library(raster)
library(lubridate)
library(adehabitatLT)
library(momentuHMM)
library(rjags)
library(bsam)
require(tidyverse)
## install aniMotum dev branch (do first time only)
# remotes::install_github("ianjonsen/aniMotum@dev")
require(aniMotum)
library(move)
library(ctmm)
library(adehabitatLT)
require(patchwork)
library(raster)
library(sp)
```

# Data pre-processing

## Raw data

Raw tracking data for Port Curtis (PC), Shoalwater Bay (SB) and Raine
Island (RI) are available upon reasonable request from the Queensland
Department of Environment, Science and Innovation
([Threatened.Species\@des.qld.gov.au](mailto:Threatened.Species@des.qld.gov.au){.email})

### Variable description

-   id: Unique identifier for turtle
-   qi: Quality index= number of source satellites for FGPS
-   DateTime: Time stamp in GMT, format YYYY-mm-dd HH:MM:SS
-   Type: ARGOS or FGPS
-   sex: Turtle sex
-   ccl: Curved carapace length
-   age.class: Maturity status determined by laparoscopy
-   lc: ARGOS location class, set to G for FGPS. The data contain ARGOS
    location classes 1,2,3 only.
-   lat: Latitude
-   lon: Longitude

Pre-processing steps:

1.  RI foraging periods were visually separated from post-nesting in
    QGIS. Removed non-foraging periods from SI and PC tracks.
2.  Removed everything before release time for PC or first 24 hours for
    SB
3.  Removed any duplicates or empty rows
4.  Combined data for PC, SB and RI = "combined"

Pre-filtering using SDLfilter improved habitat model fits:

```{r}
# data-driven filter
combined<-combined%>%mutate(Type=ifelse(lc=="G", "FGPS", "ARGOS"))%>%
  mutate(DateTime=as.POSIXct(DateTime, format='%Y-%m-%d %H:%M:%S', tz="GMT"))%>%
  group_by(id)%>%arrange(DateTime) # Make sure data are ordered by timestamp
GPS<-combined%>%filter(Type=="FGPS")%>%ungroup()
ARGOS<-combined%>%filter(Type=="ARGOS")%>%ungroup()

GPS<-as.data.frame(GPS)
ARGOS<-as.data.frame(ARGOS)


dd.GPS<-ddfilter(GPS, vmax=9.9, qi=4, ia=90, vmaxlp=2.0)
dd.ARGOS<-ddfilter(ARGOS, vmax=9.9, qi=1, ia=90, vmaxlp=2.0)


dd.data<-rbind(dd.GPS, dd.ARGOS)
```

Remove points on land

```{r}
# Load the Intertidal Extents Model Relative Extents Model (Geoscience Australia)
# https://ecat.ga.gov.au/geonetwork/srv/api/records/7d6f3432-5f93-45ee-8d6c-14b26740048a

raster_data <- raster("enviro datasets/100482_RelativeExtentsLayer/ITEM_REL_mosaic_1987_2015.tif")

# Load the point data
LatLon <- data.frame(y=dd.data$lat, x=dd.data$lon)
coordinates(LatLon) <- ~x+y
mypoints = SpatialPoints(LatLon, proj4string = CRS("+init=epsg:4326"))


# Extract the raster values at the locations of the points
myvalues = raster::extract(raster_data, mypoints)
all<-cbind(dd.data, myvalues)

# Select the points with values less than 9 (land)
filtered_point_data <- all |> filter(myvalues<9)

```

Separating tracks with gaps of more than 3 days

```{r}
# if previous timestamp is more than 3 days previous, assign a new id
Filtered_3day<-filtered_point_data%>%
  arrange(DateTime)%>%filter(!is.na(DateTime))%>%
  group_by(id1)%>%
  mutate(timediff=DateTime-lag(DateTime), # timediff is na for the first interval of each id
         id_3daygaps=ifelse(is.na(timediff), 0, ifelse(timediff>259200, 1, 0)), 
         cumsum=cumsum(id_3daygaps), #3 day splits
         id_split=paste(id1, cumsum, sep="_")) 


# retain only tracks with at least 20 locations
Filtered20plus<-Filtered_3day%>%group_by(id_split)%>%filter(n()>=20) |> droplevels()
```

# State space models

see
<https://esajournals.onlinelibrary.wiley.com/doi/full/10.1002/ecm.1470>
State space models and flexible hierachical models- observations are
imperfect measures of temporally evolving hidden states. Can be linear
or non-linear and use a variety of statistical distributions.
Assumptions: 1. state time series evolves as a Markov process (i.e. the
state at time t, depends only on the state at the previous step t-1). 2.
observations are independent of one another once we account for
dependence on the states i.e. any dependence between observations is the
result of dependence between hidden states.

# Fit SSM with Animotum

Animotum assumes GPS data are truth - but we used user-supplied
uncertainty as estimated per \# satellites: Dujon, A.M., Lindstrom, R.T.
and Hays, G.C. (2014), The accuracy of Fastloc-GPS locations and
implications for animal tracking. Methods Ecol Evol, 5: 1162-1169.
<https://doi.org/10.1111/2041-210X.12286>

```{r}
## read data, set GPS (G) to Generic Locations (GL) i.e. specifying user-supplied uncertainty 
Locs <- Filtered20plus |> ungroup() |> 
  #dplyr::select(-1) |>
  #mutate(lc = ifelse(lc == "G", "GL", lc)) |>
  dplyr::filter(!is.na(DateTime)) ##  Remove records with NA DateTime
```

```{r}
## set up table of standard deviations by # Sats - this assumes the errors are symmetric. 
GPSerr.df <- tibble(qi = 4:11, 
                        x.sd = c(937.02, 
                                 630.75, 
                                 421.24, 
                                 384.67, 
                                 159.14, 
                                 13.19, 
                                 12.45, 
                                 10.67),
                        y.sd = x.sd)
```

```{r}
## merge SD's with location data
tmp <- left_join(Locs, GPSerr.df, by = "qi") |> dplyr::select(!c(id, id1))

ids<-unique(tmp$id_split)

## use aniMotum::format_data to simplify set up for fitting SSM
d <- format_data(tmp, id="id_split", date = "DateTime", coord = c("lon","lat"))
```

The below was scripted to run all individuals in the dataset as a loop.
The code below attempts to find the optimal time-step for each
individual's track. However, the final outputs we generated by running
each individual manually. This allowed us to specify the time_step as
appropriate, and to visually examine the diagnostic plots one at a time.

```{r}
## fit SSM using loop
results <- list() # empty list

for (i in c(1)) { #length(ids))
thisturt<-d[d$id%in%ids[i],]
thisName <- unique(thisturt$id)#taking name from the dataframe
  cat("doing", thisName, "\n")
  
   # Try different time steps
  best_result <- NULL
  for (time_step in c(4,6,12,24,48,72)) {
    # Fit the SSM
    try({
    result <- aniMotum::fit_ssm(thisturt, vmax = 3, model = "mp", time.step = time_step, map = list(rho_o = factor(NA)))
})
    # Check if fitting was successful
    if (!(result$converged=="FALSE"|result$pdHess=="FALSE")) {
      best_result <- result
      break
    }
  }
  # Handle the case where no suitable time step was found
  if (is.null(best_result)) {
    message("No suitable time step found for ID:", id)
  }
  # Store the best result
  results[[i]] <- best_result
}


# save result
saveRDS(results, "20231116fit_ssm_outputALL.Rdata")
```

Plot of movement persistence over time and map

```{r}
for (i in 1:length(results)) {
  result <- results[[i]]
thisName <- unique(result$id)#taking name from the dataframe
  cat("doing", thisName, "\n")
  # Check if the result is valid
  if (!result$converged=="FALSE") {
    # Generate the persistence estimates timeseries plot
    plot(result, type = 3, pages = 1, ncol = 2)+
    plot(result, what = "predicted", type = 4, pages = 0)
    file.plot <- paste0("ssm_plots/", result$id, ".png")
    ggsave(file = file.plot)
  }
}
```

```{r}
plot(results[[1]], type = 3, pages = 1, ncol = 2) # persistence estimates as a timeseries, 4 is track plot with locations coloured by persistence

plot(results[[1]], what="predicted", type = 4, pages = 0) # plot 2-d tracks of individuals

plot(results[[1]], what="predicted", type = 1, pages = 1) # orange ribbon is 2xSE interval. type 2 (2-d timeseries plot of the track) doesn't work when there are NAs

# predicted values are state estimates at regular time intervals, orange ellipses are 95% confidence around predicted values. 

map(results[[1]],
    pages=1,
    what = "p", normalise=T)

# behavioural estimation
# plot movement persistence along the track of each animal as timeseries
p1 <- plot(results[[1]],
           type = 3,
           pages = 1,
           ncol = 2,
           normalise = TRUE,
           group = TRUE,
           pal = "Plasma") &
  ylim(0, 1) &
  theme(legend.position = "none",
        axis.text = element_text(size = 7))
p1
```

SSM validation

```{r}
plot(results[[1]], what='fitted') #fitted and predicted (gold) locations as 1-d timeseries over obs (blue), black x means failed to pass prefilter prior to fitting. larger errors (gold bands =2SE) occur in data gaps
plot(results[[1]], "p", type=2, alpha=0.1) # the above viewed as 2-d tracks


# checking whether items in the results list converged
for (i in 1:length(results)) {
  result <- results[[i]]
thisName <- unique(result$id) #taking name from the dataframe
  cat("doing", thisName, "\n")
  

  # Check if the result is valid by saving residual plots
  if (!result$converged=="FALSE") {
    res.rw <- osar(result)
    plot(res.rw, type = "ts") | plot(res.rw, type = "qq")+
    plot(res.rw, type = "acf") | plot_spacer()
    file.plot <- paste0("residual_plots/", result$id, "-residual-plots.png")
    ggsave(file = file.plot)
  }
}
```

Diagnostics are one-step ahead prediction residual, displayed as a)
time-series, b)qq-plot and c)autocorrelation functions.

a)  should show no pattern in the residuals
b)  should show normality
c)  shouldn't show autocorrelation.

We can also compare AICc statistics of various models. However, AIC can
be misleading for timeseries models and shouldn't be used as the sole
criteria.

```{r}
#res.rw$ssm[[1]]$AICc
```

To deal with poor fit, can attempt a different kind of model (e.g. crw
instead of rw).
